---
title: "Summer Temperature & Rainfall exploration"
author: "Greg Sutcliffe"
date: "28 September 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summer Temperature & Rainfall exploration

This is a Plotly chart which investigates the **mean** temperature and **total** rainfall for the summer months (defined as May to September, inclusive) over the last 50+ years.

The aim is to investigate the sort of summers Scotland gets, raleting to the age-old argument that "Scotland never has any good summers". 

## Data source

This is based on the Met Office UKCP09 daily climate dataset, cited as:
```
Met Office (2017): UKCP09: Met Office gridded land surface climate
observations - daily temperature and precipitation at 5km resolution.
Centre for Environmental Data Analysis, 2018-07-05
http://catalogue.ceda.ac.uk/uuid/319b3f878c7d4cbfbdb356e19d8061d6
```

The data shown is for a single 5km x 5km area near while the author lives.

## Good vs Bad summers

(Chart is explained on the next slide)
```{r echo=FALSE, message=FALSE, warning=FALSE}
source('./process_cp09_data.R')
ggplotly(p)
```

## Explanation

In this chart, there are two panels:

* The upper panel displays the mean temperatue per year, with a linear regression (method = `lm`) fitted
* The lower panel displays the total rainfall per year, coloured by the mean temperature of the year (cold years in blue, hot years in red)

We note two things; first, there is a clear upward trend in the mean temperature of each summer (see the regression line added to the temperature chart). This is consistent with expectations derived from climate change research.

Second, and more interesting, there is no correlation between rainfall and mean temperature for any given summer (correlation(temp,rain) => `r cor`). This is against the common view that a "bad" summer is both cold *and* wet.

## Conclusions

The data agrees with our expectations around climate change, however the expectation that rainfall and temperature would be correlated is not true. It would seem that your view of a "bad" summer should depend on whether rain or temperature is important to the activities you would persue during the summer months

The code for this document can be viewed at

* https://github.com/GregSutcliffe/summertime

This document is licensed as CC-BY-SA